---
title: "Choropleth of Heroin-Related Deaths in Connecticut, 2012-2016"
output: html_notebook
---

This notebook is a continuation of a previous analysis, in which I looked at the evolution of drug-related deaths in Connecticut from 2012-2016. Here I aim to construct a choropleth map of the counties, colored by how many heroin-related deaths occurred in that county. 

First, let's load the csv and install the necessary libraries: 

```{r}
options(warn=-1)
install.packages('rgeos')
install.packages('maptools')
install.packages('gpclib')
install.packages('RColorBrewer')

```
```{r}
df<-read.csv('county_data.csv')
df
```


Now, let's see if we can use some R GIS packages to load the county shapefile and color it based on the values as specified in the csv.

```{r}
library(rgeos)
library(maptools)
library(gpclib)
library(RColorBrewer)

```

Let's read the shapefile.
```{r}
file.exists("~/Documents/D3_Projects/datascience-projects/county_shp/nad83/countyct_37800_0000_2010_s100_census_1_shp_nad83_feet.shp")
ct_counties <- readShapeSpatial("~/Documents/D3_Projects/datascience-projects/county_shp/nad83/countyct_37800_0000_2010_s100_census_1_shp_nad83_feet.shp") 

```
```{r}
ct_counties$NAME10 = toupper(ct_counties$NAME10)
##this adds the dataframe data into the geospatial map data 
ct_counties@data = data.frame(ct_counties@data, df[match(ct_counties@data$NAME10, df$COUNTY),])
ct_counties

```
Now that we have our map data in one place, let's construct a nice color palette and then plot the map! 

```{r}
my.palette <- brewer.pal(n = 7, name = "OrRd")
spplot(ct_counties, "NUMBER.1", col.regions=my.palette, cuts = 6, main="Heroin-related Deaths by County (2012-2016)")
##export as svg
jpeg('map.jpg')
spplot(ct_counties, "NUMBER.1", col.regions=my.palette, cuts = 6, main="Heroin-related Deaths by County (2012-2016)")
dev.off()

```